---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit the .Rmd file. -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```


<!-- badges: start -->
[![CRAN status](https://www.r-pkg.org/badges/version/ReboundTools)](https://cran.r-project.org/package=ECCTools)
[![lifecycle](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![Project Status: Active – The project has reached a stable, usable state and is being actively developed.](https://www.repostatus.org/badges/latest/active.svg)](https://www.repostatus.org/#active)
[![R-CMD-check](https://github.com/earamendia/ECCTools/workflows/R-CMD-check/badge.svg)](https://github.com/earamendia/ECCTools/actions)
[![codecov](https://codecov.io/gh/earamendia/ECCTools/branch/develop/graph/badge.svg?token=IG68VLDYEY)](https://codecov.io/gh/earamendia/ECCTools)
[![Codecov test coverage](https://codecov.io/gh/earamendia/ECCTools/branch/master/graph/badge.svg)](https://codecov.io/gh/earamendia/ECCTools?branch=master)
<!-- badges: end -->


# ECCTools

The `R` package `ECCTools` provides tools to modify and Energy Conversion Chain (ECC) 
that has been previously loaded using the `R` package `IEATools`. 
Particularly, it helps to prepare the formulation of Multi-Regional Physical Supply Use Tables (MR-PSUT)
representing the ECC, as well as to formulate Domestic Technology Assumption (DTA) formulations of national ECC.


## Installation

You can install `ECCTools` from github with:

```{r gh-installation, eval = FALSE}
# install devtools if not already installed
# install.packages("devtools")
devtools::install_github("earamendia/ECCTools")
# To build vignettes locally, use
devtools::install_github("earamendia/ECCTools", build_vignettes = TRUE)
```


## History

The paper that demonstrates uses of this package is currently under review.


## More Information

Find more information, including vignettes and function documentation, at 
<https://earamendia.github.io/ECCTools/>.

